```{r init, echo=FALSE, message=FALSE, results='hide'}
library("markdown")
options(rstudio.markdownToHTML = 
  function(inputFile, outputFile) {      
    require(markdown)
    markdownToHTML(inputFile, outputFile, stylesheet='custom.css')   
  }
)
```

Analisys of the Leap Motion experiment at UCP
========================================================


This document provides the analysis of the experiment conducted at UCP/CITAR comparing the movement of three devices - Mouse, Touchpad, and Leap Motion (via Touchless app) - for pointing tasks. 

We use Maczenzie's accuracy metrics to compare the devices.

```{r setup, message=FALSE}

PRODUCE_BULK_CHARTS = FALSE
PRODUCE_BULK_HIST_CHARTS = TRUE
PRODUCE_BULK_CHARTS_DEVICE_PLUS_USERID = FALSE

#install.packages(c("ggplot2", "doBy", "psych", "car", "xtable"))
require(ggplot2)
require(doBy)
require(psych)
require(car)
require(xtable)
source("functions.R")

# load metrics data
data <- read.csv(file="data/measures-feup.txt", head=TRUE, sep="")

# Users 0 and 5 did not complete the experiment
data <- data[ data$UserId != 5,]

# Only blocks <= 7
data <- data[ data$Block <= 7,]

# convert the column to factor and name the levels
data$Device <- as.factor(data$Device)

# Set the levels 
# LeapMotion=c(0),  Mouse=c(1), Touchpad=c(2), LeapMotionTouchless=c(4)
# LeapMotion does not exist in the dataset from the UCP experiment
levels(data$Device) <- list(Mouse=c(1), LeapMotion=c(0), LeapMotionTouchless=c(4))

# convert the block column to factor
data$Block <- factor(data$Block)

# In this analysis we use ErrorRateBinary, so we drop the ErrorRate column and change the column name ErrorRateBinary to ErrorRate
drops <- c("ErrorRate")
data<-data[,!(names(data) %in% drops)]
names(data)[names(data) == "ErrorRateBinary"] <- "ErrorRate"

# Compute the aggregate means for each variable.
# We agregate all values for the various circles
IVars <- c("Device", "Block", "Sequence", "UserId")
DVars <- c("Throughput", "ErrorRate", "MovementTime", "TRE", "TAC", "ODC", "MDC", "MV", "ME", "MO")
aggData <-aggregate(data[ ,DVars], data[ ,IVars], mean)
aggData$Block <- factor(aggData$Block)


```

Movement time as function of block
-------------------------------------
```{r movement-time-block, fig.width=21/2.54, cache=TRUE}

describe.MT.aggData <- describeBy(aggData$MovementTime, aggData[,c("Block")])
describe.MT.aggData <-do.call("rbind", describe.MT.aggData)
describe.MT.aggData <- cbind(Block=rownames(describe.MT.aggData), describe.MT.aggData)
describe.MT.aggData$Block <- factor(describe.MT.aggData$Block)

p <- ggplot(describe.MT.aggData, aes(x=Block, y=mean )) + 
    geom_point(stat="identity") +
    geom_line(aes(group=vars)) + 
    ylab("Movement time (seconds) with 95% conf int") +
    #xlab("Sequence") +
    #ggtitle("Throughput") +
    geom_errorbar(aes(ymin=mean-1.96*se, ymax=mean+1.96*se), colour="Black",
                  width=.2                    # Width of the error bars
                  ) +
    theme(legend.position="none") +
    theme() 
p

p <- ggplot(aggData, aes_string(x="Block", y="MovementTime", group="Device", colour="Device" )) + 
            stat_summary(fun.y="mean", geom="line") + 
            stat_summary(fun.y="mean", geom="point", aes(shape=Device)) + 
            theme(legend.direction = "horizontal", legend.position = "top") +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            scale_fill_brewer(palette="Set1") + 
            scale_colour_brewer(palette="Set1") +
            theme() #noop
p
ggsave(file = paste("charts/","all-movementtime-block-lineplot.pdf", sep=""), 
               width=13/2.54, height=7/2.54, dpi=100)
```

<!-- a comment here 
```{r movement-time-block-peruser-leapmotiontouchless, fig.width=21/2.54, fig.height=40/2.54, results='hide', message=FALSE, echo=FALSE, fig.show='hide', cache=TRUE}
p <- ggplot(aggData[aggData$Device=="LeapMotionTouchless",], aes_string(x="Block", y="MovementTime", group="UserId", colour="UserId" )) + 
            stat_summary(fun.y="mean", geom="line") + 
            stat_summary(fun.y="mean", geom="point", aes(shape=Device)) + 
            theme(legend.direction = "horizontal", legend.position = "top") +
            facet_wrap( ~ UserId, ncol=2) +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            theme() #noop
p
```
-->

Throughput  as function of block
-------------------------------------
```{r throughput-block, fig.width=21/2.54, cache=TRUE}


describe.T.aggData <- describeBy(aggData$Throughput, aggData[,c("Block")])
describe.T.aggData <-do.call("rbind", describe.T.aggData)
describe.T.aggData <- cbind(Block=rownames(describe.T.aggData), describe.T.aggData)
describe.T.aggData$Block <- factor(describe.T.aggData$Block)

p <- ggplot(describe.T.aggData, aes(x=Block, y=mean )) + 
    geom_point(stat="identity") +
    geom_line(aes(group=vars)) + 
    ylab("Throughput (bps) with 95% conf int") +
    #xlab("Sequence") +
    #ggtitle("Throughput") +
    geom_errorbar(aes(ymin=mean-1.96*se, ymax=mean+1.96*se), colour="Black",
                  width=.2                    # Width of the error bars
                  ) +
    theme(legend.position="none") +
    theme() 
p


p <- ggplot(aggData, aes_string(x="Block", y="Throughput", group="Device", colour="Device" )) + 
            stat_summary(fun.y="mean", geom="line") + 
            stat_summary(fun.y="mean", geom="point", aes(shape=Device)) + 
            theme(legend.direction = "horizontal", legend.position = "top") +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            theme() #noop
p
ggsave(file = paste("charts/","all-throughput-block-lineplot.pdf", sep=""), 
               width=20/2.54, height=16/2.54, dpi=100)
```

Error rate  as function of block
-------------------------------------
```{r ErrorRate-block, fig.width=21/2.54, cache=TRUE}
p <- ggplot(aggData, aes_string(x="Block", y="ErrorRate", group="Device", colour="Device" )) + 
            stat_summary(fun.y="mean", geom="line") + 
            stat_summary(fun.y="mean", geom="point", aes(shape=Device)) + 
            theme(legend.direction = "horizontal", legend.position = "top") +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            theme() #noop
p
ggsave(file = paste("charts/","all-throughput-block-lineplot.pdf", sep=""), 
               width=20/2.54, height=16/2.54, dpi=100)
```

Learning effect
-------------------------
To estimate the learning effect, we ran pairwise t-tests for average throughput per block (considering all devices) with a significance level of 5%:

```{r learning-effect}
#describeBy(data$MovementTime, data$Block)

# pairwise t-tests to determine learning effect blocks
p.t.test <- pairwise.t.test(aggData$Throughput, aggData$Block, paired=T, p.adjust.method="none")
diag(p.t.test$p.value) 
diag(p.t.test$p.value) < 0.05


# We will consider only blocks 4 to 8 in the rest of the analysis
filenameprefix <- "blocks4-7"
aggData.noLearn <- aggData[as.numeric(aggData$Block) > 3,]

# Drop unused block levels
aggData.noLearn$Block <- factor(aggData.noLearn$Block)
```

The results indicate a clear learning effect in blocks 1 to 3, but also indicate a significant different between blocks 6 and 7, suggesting that participants were still learning after block 6. However, in our following analysis we discard only blocks 1 to 3, since those represent the most significant learning effect.

```{r learning-effect-other, echo=FALSE, message=FALSE, results='hide'}
# pairwise t-tests to determine learning effect blocks on Movement time
for ( device in unique(data$Device)) {
   p = pairwise.t.test(data$MovementTime[data$Device==device], data$Block[data$Device==device], 
        paired=T, p.adjust.method="none")
  print(device)
  print(p)
}

# pairwise t-tests to determine learning effect blocks on Throughput
for ( device in unique(data$Device)) {
   p = pairwise.t.test(data$Throughput[data$Device==device], data$Block[data$Device==device], paired=T, p.adjust.method="none")
    print(device)
    print(p)
}


```



Summary statistics 
------------

```{r summary-statistics, results='asis'}
options(width = 200)

# describe the main variables and store in file
s<-describeBy(aggData.noLearn[, DVars], aggData.noLearn$Device, mat=FALSE, digits=2)
s <- do.call("rbind", s)

print(xtable(s), type = "html")

s <- cbind(Device=rownames(s), s)

write.table(s, file = paste("tables/", filenameprefix,"-measures-device-describeby.csv", sep=""), sep=",", row.names=FALSE)

remove(s)


# descriptive stats for device*userid
s1<-describeBy(aggData.noLearn[,DVars], 
               list(aggData.noLearn$Device,aggData.noLearn$UserId), mat=FALSE, digits=2)
s1 <- do.call("rbind", s1)
s1 <- cbind(Device=rownames(s1), s1)

write.table(s1, file = paste("tables/", filenameprefix,"-measures-device+userid-describeby.csv", sep=""), sep=",", row.names=FALSE)

remove(s1)
#s1
```

Histograms
-----------------
```{r charts-bulk-hist, echo=FALSE, message=FALSE, fig.width=21/2.54, cache=TRUE}
if (PRODUCE_BULK_HIST_CHARTS == TRUE) {
    for (var in DVars ) {
        # plot the histograms of the for each device
        p <- ggplot(aggData.noLearn, aes_string(x=var, colour="Device" )) + 
            geom_histogram() + 
            #coord_flip() + 
            facet_wrap( ~ Device, ncol=1) +
            theme(legend.position="none") +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            theme()
        print(p)
        ggsave(file = paste("charts/hist/",filenameprefix,"-",var,"-hist.pdf", sep=""), 
               width=20/2.54, height=16/2.54, dpi=100)
        
        if (PRODUCE_BULK_CHARTS_DEVICE_PLUS_USERID == TRUE) {
            # plot the histograms of the movement time for each device
            p <- ggplot(aggData.noLearn, aes_string(x=var, colour="Device" )) + 
                geom_histogram() + 
                #coord_flip() + 
                facet_wrap(Device ~ UserId, nrow=3) +
                theme(legend.position="none") +
                theme()
            print(p)
            ggsave(file = paste("charts/hist/",filenameprefix,"-",var,"-device+userid-hist.pdf", sep=""), 
                width=20/2.54, height=16/2.54, dpi=100)
        }
  
  
    }
}

```

Boxplots
----------------
```{r charts-bulk-boxplot, echo=FALSE, message=FALSE}
if (PRODUCE_BULK_CHARTS == TRUE) {
    for (var in DVars ) {
        # plot the boxplots of the movement time for each device
        p <- ggplot(aggData.noLearn, aes_string(x="Device", y=var,  colour="Device" )) + 
            geom_boxplot() + 
            coord_flip() + 
            theme(legend.position="none") +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            theme()
        print(p)
        ggsave(file = paste("charts/boxplot/",filenameprefix,"-",var,"-boxplot.pdf", sep=""), 
               width=20/2.54, height=16/2.54, dpi=100)
        
        if ( PRODUCE_BULK_CHARTS_DEVICE_PLUS_USERID == TRUE ) {
            # plot the boxplots of the movement time for each device
            p <- ggplot(aggData.noLearn, aes_string(x="Device", y=var,  colour="Device" )) + 
                geom_boxplot() + 
                #coord_flip() + 
                facet_wrap( ~ UserId, nrow=1) +
                theme(legend.position="none") +
                theme()
            print(p)
            ggsave(file = paste("charts/boxplot/",filenameprefix,"-",var,"-device+userid-boxplot.pdf", sep=""), 
               width=20/2.54, height=16/2.54, dpi=100)
         }
    }
}
```


Lineplots
-------------
```{r charts-bulk-lineplot, echo=FALSE, message=FALSE}
if (PRODUCE_BULK_CHARTS == TRUE) {
    for (var in DVars ) {
        # line plot movement time over block for each device
        p <- ggplot(aggData.noLearn, aes_string(x="Block", y=var, group="Device", colour="Device" )) + 
            stat_summary(fun.y="mean", geom="line") + 
            stat_summary(fun.y="mean", geom="point", aes(shape=Device)) + 
            theme(legend.direction = "horizontal", legend.position = "top") +
            #ylab("Percentage (%)") +
            #xlab("Sequence") +
            #ggtitle(var) +
            theme() #noop
        print(p)
        ggsave(file = paste("charts/lineplot/",filenameprefix,"-",var,"-block-lineplot.pdf", sep=""), 
               width=20/2.54, height=16/2.54, dpi=100)
        
        
        if ( PRODUCE_BULK_CHARTS_DEVICE_PLUS_USERID == TRUE ) {
  
            # line plot movement time over block for each device
            p <- ggplot(aggData.noLearn, aes_string(x="Block", y=var, group="Device", colour="Device" )) + 
                stat_summary(fun.y="mean", geom="line") + 
                stat_summary(fun.y="mean", geom="point", aes(shape=Device)) + 
                facet_wrap( ~ UserId, nrow=3) +
                theme(legend.direction = "horizontal", legend.position = "top") +
                theme() #noop
            p
            ggsave(file = paste("charts/lineplot/",filenameprefix,"-",var,"-device+userid-block-lineplot.pdf", 
                                sep=""), width=20/2.54, height=16/2.54, dpi=100)
        }
  
    }
}
```

MO Over TRE
-----------------------
```{r chart-MO-over-TRE}
s <- summaryBy(ErrorRate+TRE+TAC+MDC+ODC+MV+ME+MO+MovementTime+Throughput~Device, 
               data=aggData.noLearn,
               FUN=c(mean,sd))
# plot the movement offset over the TRE for each device
p <- ggplot(s, aes(x=TRE.mean, y=abs(MO.mean), group=Device, colour=Device )) +
    geom_point(size=5, aes(shape=Device)) + 
    #coord_cartesian(xlim = c(0, 0.4), ylim=c(0, 8)) + 
    ylab("Movement offset (pixels)") +
    xlab("Target Re-Entry") +
    theme(legend.position=c(0.55,0.9), legend.direction="horizontal") +
    theme()
p
ggsave(file = paste("charts/",filenameprefix,"-MO-TRE.pdf", sep=""), width=14/2.54, height=7/2.54, dpi=100)



```

Throughput and Error rate
-----------------
```{r chart-Throughput-and-ErrorRate, fig.show='hold', message=FALSE, warning=FALSE}
# plot the throughput and ErrorRate

describe <- describeBy(aggData.noLearn$Throughput, list(aggData.noLearn$Device))
describe<-do.call("rbind",describe)
describe <- cbind(Device=rownames(describe), describe)
levels(describe$Device) <- list( Mouse=c("Mouse"), LeapMotion=c("LeapMotion"), LeapMotionTouchless=c("LeapMotionTouchless"))
p1 <- ggplot(describe, aes(x=Device, y=mean, colour=Device, fill=Device )) + 
    geom_bar(stat="identity") +
    ylab("BPS") +
    xlab("Throughput") +
    #ggtitle("Throughput") +
    geom_errorbar(aes(ymin=mean-1.96*se, ymax=mean+1.96*se), colour="Black",
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
    theme(legend.position="none", 
          #axis.text.x = element_text(angle = 20, hjust = 1),
          axis.text.x=element_blank()) +
    scale_fill_brewer(palette="Set1") + 
    scale_colour_brewer(palette="Set1") +
    theme() 

#ggsave(file = paste("charts/",filenameprefix,"-Throughput.pdf",sep=""), width=20/2.54, height=16/2.54, dpi=100)

describe <- describeBy(aggData.noLearn$ErrorRate, list(aggData.noLearn$Device))
describe<-do.call("rbind",describe)
describe <- cbind(Device=rownames(describe), describe)
levels(describe$Device) <- list( Mouse=c("Mouse"), LeapMotion=c("LeapMotion"), LeapMotionTouchless=c("LeapMotionTouchless"))

p2 <- ggplot(describe, aes(x=Device, y=mean*100, colour=Device, fill=Device )) + 
    geom_bar(stat="identity") +
    ylab("Percent.(%)") +
    xlab("ErrorRate") +
    #ggtitle("ErrorRate") +
    geom_errorbar(aes(ymin=100*(mean-1.96*se), ymax=100*(mean+1.96*se)), colour="Black",
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9)) +
    
    theme(legend.position=c(0,0.9), legend.direction="horizontal", 
          #axis.text.x = element_text(angle = 20, hjust = 1),
          axis.text.x=element_blank()) +
    scale_fill_brewer(palette="Set1") + 
    scale_colour_brewer(palette="Set1") +
    theme() 

#pdf(file = paste("charts/",filenameprefix,"-Throughput+ErrorRate.pdf",sep=""), width=20/2.54, height=16/2.54)
pdf(file = paste("charts/",filenameprefix,"-Throughput+ErrorRate-small.pdf",sep=""), width=16/2.54, height=7/2.54)
multiplot(p1, p2, cols=2)
dev.off()

p1
p2
```


Barplots for metrics
-----------------
```{r chart-barplots, fig.show='hold', message=FALSE, warning=FALSE}

describeMetrics <- describeBy(aggData.noLearn[, c("TRE","TAC","MDC","ODC","MV","ME","MO")], list(aggData.noLearn$Device))
a=cbind(describeMetrics$LeapMotionTouchless, metric=rownames(describeMetrics$LeapMotionTouchless))
a=cbind(a, Device="LeapMotionTouchless")
b=cbind(describeMetrics$Mouse, metric=rownames(describeMetrics$Mouse))
b=cbind(b, Device="Mouse")
c=cbind(describeMetrics$Touchpad, metric=rownames(describeMetrics$Touchpad))
c=cbind(c, Device="Touchpad")
describeMetrics = rbind(a, b, c)
describeMetrics$metric <- factor(describeMetrics$metric, levels=c("TRE", "TAC", "MDC", "ODC", "MV", "ME", "MO"))

levels(describeMetrics$Device) <- list( Mouse=c("Mouse"), Touchpad=c("Touchpad"), LeapMotionTouchless=c("LeapMotionTouchless"))

ggplot(describeMetrics, aes(x=Device, y=abs(mean), group=Device, colour=Device, fill=Device)) + 
    #stat_summary(fun.y="mean", geom="bar") + 
    geom_bar(stat="identity", width=.5, position = position_dodge(width=0.5)) +
    #geom_bar() +
    #coord_flip() +
    ylab("Mean") +
    xlab("") +
    geom_errorbar(aes(ymin=abs(mean)-1.96*abs(se), ymax=abs(mean)+1.96*abs(se)), colour="Black",
                  width=.2,                    # Width of the error bars
                  size = .1,
                  position=position_dodge(.5)) +
    facet_wrap(  ~ metric,nrow=1, scales="free") +
    theme(legend.position=c(.5,-0.1), legend.direction="horizontal", 
          axis.text.x = element_blank()) + 
    scale_fill_brewer(palette="Set1") + 
    scale_colour_brewer(palette="Set1") 
ggsave(file = paste("charts/",filenameprefix,"-metrics.pdf", sep=""), width=30/2.54, height=6/2.54, dpi=100)

```

Correlations
---------------
```{r correlations, results='asis'}
vars <- c("Throughput", "TRE", "TAC", "ODC", "MDC", "MV", "ME", "MO")


print(xtable(round(cor(aggData.noLearn[, vars]), 2)), type = "html")


```


Analysis of Variance - Differences between devices
---------------------
Analysis of variance of the throughput for the three devices showed significant differences 
```{r anova}

#ggplot(aggData.noLearn, aes(x=1:nrow(aggData.noLearn), y=MO, group=Device, colour=Device)) + geom_point()

#ggplot(aggData.noLearn, aes(x=Throughput, y=MO,  colour=Device)) + geom_point()



leveneTest(Throughput ~ Device,data=aggData.noLearn, center="mean")
leveneTest(ErrorRate ~ Device,data=aggData.noLearn, center="mean")
leveneTest(TAC ~ Device,data=aggData.noLearn, center="mean")
leveneTest(TRE ~ Device,data=aggData.noLearn, center="mean")
leveneTest(MDC ~ Device,data=aggData.noLearn, center="mean")
leveneTest(ODC ~ Device,data=aggData.noLearn, center="mean")
leveneTest(ME ~ Device,data=aggData.noLearn, center="mean")
leveneTest(MV ~ Device,data=aggData.noLearn, center="mean")
leveneTest(MO ~ Device,data=aggData.noLearn, center="mean")

for ( var in DVars ) {
    print(paste("Variances for ", var))
    varMouse = var(aggData.noLearn[aggData.noLearn$Device=="Mouse", var])
    varTouch = var(aggData.noLearn[aggData.noLearn$Device=="Touchpad", var])
    varLeap = var(aggData.noLearn[aggData.noLearn$Device=="LeapMotionTouchless", var])
    all = c(varMouse, varTouch, varLeap)
    print (paste("Mouse: ", varMouse, " Touchpad: ", varTouch, " LeapMotionTouchless: ", varLeap, " Max/Min: ", max(all)/min(all)))
} 


# Anova for Throughput
aov.Throughput <- aov(Throughput~Device+Error(factor(UserId) / Device), data=aggData.noLearn)
#aov.Throughput <- aov(Throughput~Device, data=aggData.noLearn)
#aov.Throughput
summary(aov.Throughput)

# pairwise comparisons, adjusting p-value
# all devices are significantly different from each other, in terms of throughput
TukeyHSD(aov.Throughput)



# Anova for TAC
aov.TAC <- aov(TAC~Device, data=aggData.noLearn)
#aov.TAC
summary(aov.TAC)
TukeyHSD(aov.TAC)



# Anova for TRE
aov.TRE <- aov(TRE~Device, data=aggData.noLearn)
#aov.TRE
summary(aov.TRE)
TukeyHSD(aov.TRE)


# Anova for MDC
aov.MDC <- aov(MDC~Device, data=aggData.noLearn)
#aov.MDC
summary(aov.MDC)
TukeyHSD(aov.MDC)


# Anova for ODC
aov.ODC <- aov(ODC~Device, data=aggData.noLearn)
#aov.ODC
summary(aov.ODC)
TukeyHSD(aov.ODC)

# pairwise.t.test(aggData.noLearn$ODC, aggData.noLearn$Device,  paired=F, pool=F)
# t.test(aggData.noLearn$ODC[aggData.noLearn$Device == "Mouse"], aggData.noLearn$ODC[aggData.noLearn$Device == "Touchpad"],  paired=F, var.equal=T)

# Anova for MV
aov.MV <- aov(MV~Device, data=aggData.noLearn)
#aov.MV
summary(aov.MV)
TukeyHSD(aov.MV)


# Anova for ME
aov.ME <- aov(ME~Device, data=aggData.noLearn)
#aov.ME
summary(aov.ME)
TukeyHSD(aov.ME)


# Anova for MO
aov.MO <- aov(MO~Device, data=aggData.noLearn)
#aov.MO
summary(aov.MO)
TukeyHSD(aov.MO)
```

```{r lm}

model.tre <- lm(Throughput~TRE, data=aggData.noLearn)
summary(model.tre)


model.odc <- lm(Throughput~ODC, data=aggData.noLearn)
summary(model.odc)

model.mdc <- lm(Throughput~MDC, data=aggData.noLearn)
summary(model.mdc)

model.odcmdc <- lm(Throughput~ODC+MDC, data=aggData.noLearn)
summary(model.odcmdc)

model.treodcmdc <- lm(MovementTime~TRE+ODC+MDC, data=aggData.noLearn)
summary(model.treodcmdc)

model.tretac <- lm(Throughput~TRE+TAC, data=aggData.noLearn)
summary(model.tretac)

```


MDC Over ODC
-----------------------
```{r chart-MDC-over-ODC}
s <- summaryBy(ErrorRate+TRE+TAC+MDC+ODC+MV+ME+MO+MovementTime+Throughput~Device, 
               data=aggData.noLearn,
               FUN=c(mean,sd))
# plot the movement offset over the TRE for each device
p <- ggplot(s, aes(x=MDC.mean, y=abs(ODC.mean), group=Device, colour=Device )) +
    geom_point(size=5, aes(shape=Device)) + 
    #coord_cartesian(xlim = c(0, 0.4), ylim=c(0, 8)) + 
    ylab("Orthogonal direction change (pixels)") +
    xlab("Movement direction change (pixels)") +
    theme(legend.position=c(0.55,0.9), legend.direction="horizontal") +
    theme()
p
ggsave(file = paste("charts/",filenameprefix,"-MDC-ODC.pdf", sep=""), width=14/2.54, height=7/2.54, dpi=100)



```